Improving the communication and transparency of stock assessment using interactive

### 2018-12-11 analysis/paper/manuscript.Rmd
Tweak ms and start producing webshots of the NCAM dashboard

In the age of data-intensive science and big data, scientists across many fields are faced with the challenge of synthesizing and communicating information from large and complex data sets. Scientific data are often presented as tabular summaries and static plots, however, seeing patterns in tables can be difficult and static plots preclude detailed explorations of the underlying data. New open source tools are bridging this gap by making it easier to build interactive figures [@perkel2018]. In particular, browser-based data visualizations have been playing an increasingly prominent role in communicating information on a wide range of topics. Interactive graphics are frequently utilized by media outlets, such as the FiveThirtyEight, and scientists from a wide range of disciplines are starting to apply interactive visualization tools to explore and communicate their results [@yeatman2018; @tushar2017; @wick2015]. This trend has been fueled by the development of JavaScript libraries (e.g. D^3^ [@bostock2011], plotly [@plotly], leaflet [@leaflet], highcharts [@highcharts]) that enable the production of dynamic and interactive data visualizations in modern web-browsers. Interfaces to JavaScript libraries have also been developed to allow interactive plots to be coded using commonly used languages such as R [@R] or Python [@python]. By laying the technical infrastructure for creating interactive graphics, these new tools are ushering in a new era of Big Data by facilitating the exploration and communication of complex data and models.

@@

Synthesizing data from multiple sources presents a key challenge to stock assessment. Analyses of different data sources were traditionally carried out independently and the summaries or parameters from these analyses were used in the assessment model. This approach, however, is less than ideal because information may be lost and uncertainty may be unaccounted for when we "do statistics on the statistics" [@link1999; @maunder2013]. Such issues have largely been curtailed in contemporary stock assessments thanks to advances in software that have facilitated the analysis of all available data, in as raw a form as appropriate, in a single integrated analysis [@maunder2013]. Specifically, statistical modeling tools such as JAGS [@plummer2003], AD Model Builder [@fournier2012] and Template Model Builder [@kristensen2015] allow the construction of a joint likelihood for an array of observations to, in theory, extract as much information as possible about the biological and fishery processes. However, integrated analyses are not a panacea because model misspecifications and data conflicts are an inevitable consequence of simplifying reality to a small series of equations [@maunder2017]. A potential solution to this quandary to use a superensemble model, whereby multiple models with different structures are run and their predictions are supplied as covariates to an additional superensemble model [@anderson2017]. Ensemble approaches reduce the risk of picking the "wrong" model and also expands the range of hypotheses explored [@dietterich2000]. These advances greatly improve our ability to assess the status and trends of fish populations, however, modern stock assessment biologists are now faced with the overwhelming task of understanding an ever expanding array of data inputs and model outputs. This challenge is compounded by the ever increasing demand for rapid and sound fisheries advice. Here we argue that the process of providing fisheries advice can be further streamlined by utilizing dynamic and interactive visualization tools to explore, diagnose and communicate stock assessment models. We hope to demonstrate this using a recently developed interactive tool for exploring an integrated assessment model for Northern cod.

The Northern cod stock off southern Labrador and eastern Newfoundland is one of the most well studied stocks in eastern Canada, perhaps by virtue of its history. As such, there are multiple monitoring programs that help inform the status of the stock and data from most of these programs have been integrated into a state-space stock assessment model, called NCAM [@cadigan2016]. The "base case" model includes information from research vessel autumn trawl surveys (1983-present), Sentinel fishery surveys (1995-present), inshore acoustic surveys (1995-2009), fishery catch-at-age compositions and partial fishery landings (1983-present), and tagging (1983-present). Using a series of observation equations, this TMB based model reduces thousands of historical data points from these monitoring programs into quantities such as recruitment, spawning stock biomass, fishing mortality and natural mortality. Once the model is fit to the data, the next step is to produce visual representations of the data and model. An interactive and self-contained "dashboard" was developed for this purpose for the last assessment of Northern cod (**TODO: add reference**). The concept of using a dashboard was burrowed from the buisness community where dashboards are frequently used to group a series of interactive visuals and tables to provide at-a-glance views of key performance indicators. Many tools have been developed to facilitate the development of data dashboards and we leveraged R-based tools to construct a tool for exploring the input and output of NCAM. Specifically we utilized the flexdashboard [@allaire2017] package to group interactive plotly-based [@plotly] visuals into a dynamic document. We also used the crosstalk [@cheng2016] package to link the data displayed across multiple plots in the "Tagging" tab.

TODO:
- Take screenshots of the dashboard for static figures for the paper (use webshot?)
- Remember to add reference to the supplement that will ultimatly include the NCAM_dashboard.html

>>>>

In the age of data-intensive science and big data, scientists across many fields are faced with the challenge of synthesizing and communicating information from large and complex data sets. Scientific data are often presented as tabular summaries and static plots, however, seeing patterns in tables can be difficult and static plots preclude detailed explorations of the underlying data. New open source tools are bridging this gap by making it easier to build interactive figures [@perkel2018]. In particular, browser-based data visualizations have been playing an increasingly prominent role in communicating information on a wide range of topics. Interactive graphics are frequently utilized by media outlets, such as the FiveThirtyEight, and scientists from a wide range of disciplines are starting to apply interactive visualization tools to explore and communicate their results [@yeatman2018; @tushar2017; @wick2015]. This trend has been fueled by the development of JavaScript libraries (e.g. D^3^ [@bostock2011], plotly [@plotly], leaflet [@leaflet], highcharts [@highcharts]) that enable the production of dynamic and interactive data visualizations in modern web-browsers. Interfaces to JavaScript libraries have also been developed to allow interactive plots to be coded using commonly used languages such as R [@R] or Python [@python]. By laying the technical infrastructure for creating interactive graphics, these new tools are ushering in a new era of big data by facilitating the exploration and communication of complex data and models.

@@

Synthesizing data from multiple sources presents a key challenge to stock assessment. Analyses of different data sources were traditionally carried out independently and the summaries or parameters from these analyses were used in the assessment model. This approach, however, is less than ideal because information may be lost and uncertainty may be unaccounted for when we "do statistics on the statistics" [@link1999; @maunder2013]. Such issues have largely been curtailed in contemporary stock assessments thanks to advances in software that have facilitated the analysis of all available data, in as raw a form as appropriate, in a single integrated analysis [@maunder2013]. Specifically, statistical modeling tools such as JAGS [@plummer2003], AD Model Builder [@fournier2012] and Template Model Builder [@kristensen2015] allow the construction of a joint likelihood for an array of observations to, in theory, extract as much information as possible about the biological and fishery processes. However, integrated analyses are not a panacea because model misspecifications and data conflicts are an inevitable consequence of simplifying reality to a small series of equations [@maunder2017]. A potential solution to this quandary to use a superensemble model, whereby multiple models with different structures are run and their predictions are supplied as covariates to an additional superensemble model [@anderson2017]. Ensemble approaches reduce the risk of picking the "wrong" model and also expands the range of hypotheses explored [@dietterich2000]. These advances greatly improve our ability to assess the status and trends of fish populations, however, modern stock assessment biologists are now faced with the overwhelming task of understanding an ever expanding array of data inputs, model structures and outputs. This challenge is compounded by the ever increasing demand for rapid and sound fisheries advice. Here we argue that the process of providing fisheries advice can be further streamlined by utilizing dynamic and interactive visualization tools to explore, diagnose and communicate stock assessment models. We hope to demonstrate this using a recently developed interactive tool for exploring an integrated assessment model for Northern cod.

The Northern cod stock off southern Labrador and eastern Newfoundland is one of the most well studied stocks in eastern Canada, perhaps by virtue of its history. As such, there are multiple monitoring programs that help inform the status of the stock and data from most of these programs have been integrated into a state-space stock assessment model, called NCAM [@cadigan2016]. The "base case" model includes information from research vessel autumn trawl surveys (1983-present), Sentinel fishery surveys (1995-present), inshore acoustic surveys (1995-2009), fishery catch-at-age compositions and partial fishery landings (1983-present), and tagging (1983-present). Using a series of observation equations, this TMB based model reduces thousands of historical data points from these monitoring programs into quantities such as recruitment, spawning stock biomass, fishing mortality and natural mortality. Once the model is fit to the data, the next step is to produce visual representations of the data and model. An interactive and self-contained "dashboard" was developed for this purpose for the last assessment of Northern cod [@dwyer2019]. The concept of using a dashboard was burrowed from the business community where dashboards are frequently used to group a series of interactive visuals and tables to provide at-a-glance views of key performance indicators. Many tools have been developed to facilitate the development of data dashboards and we leveraged R-based tools to construct a tool for exploring the input and output of NCAM. Specifically we utilized the flexdashboard [@allaire2017] package to group interactive plotly-based [@plotly] visuals into a dynamic document. We also used the crosstalk [@cheng2016] package to link the data displayed across multiple plots in the "Tagging" tab.

TODO:
- Take screenshots of the dashboard for static figures for the paper (use webshot?)
- Remember to add reference to the supplement that will ultimately include the NCAM_dashboard.html
